\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2018-01-18 eg6-letters/NIM/EG6_RTPC_NIM.tex
Improving the NIM paper

\documentclass[preprint,5p]{elsarticle}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}  
\usepackage{amssymb}  
\usepackage{hyperref}
\providecommand{\keywords}[1]{\textbf{\textit{Keywords:}} #1}
\hypersetup{colorlinks=true, urlcolor=blue, citecolor=blue}
\usepackage[displaymath]{lineno}
\linenumbers
\usepackage{dblfloatfix}    % To enable figures at the bottom of page
\begin{document}

@@

The original design of the RTPC was developed for the BoNuS 
experiment at Jefferson Lab which took data with CLAS in 
2005~\cite{Fenker:2008zz}. Significant improvements were made to the RTPC mechanical 
structure and fabrication technique that both increased the acceptance and 
reduced the amount of material in the path of the outgoing particles. The enhanced design, 
used in the 2009 DVCS experiment, is described in section \ref{sec_design} of 
this paper. The data acquisition system described in section \ref{sec_readout} 
was reconfigured to increase the event readout rate. The calibration methods 
are discussed in section \ref{sec_calib} and the tracking algorithm in section
\ref{sec_tracking}. Finally, the overall performance
of the RTPC is described in section \ref{sec_perfor}.

@@

We detail here the different regions shown in Figure~\ref{fig:RTPC_1_4} 
starting from the beam line towards larger radius:\\
\begin{itemize}
  \item The 6~atm $^4$He target extends along the beamline forming the detector 
     central axis; it is a 292~mm~long, 6~mm diameter Kapton straw with a 
     27~$\mu$m wall. The detector and the target are centered at 64~cm upstream 
     of the CLAS center.
   \item The first gas gap covers the radial range from 3~mm to 20~mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 $\mu$m thick window made of grounded aluminized Mylar.
   \item The second gas gap region extends between 20~mm and 30~mm and is 
      filled with the gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether 
      (DME).  This region is surrounded by a 4~$\mu$m thick window made of 
      aluminized Mylar set at $-4260$~V to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and 
      extends from the cathode to the first GEM, 60~mm away from the beam axis.  
      The average electric field in this region is perpendicular to the beam 
      and about 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66~mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600~V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charges 
      after they have been multiplied by the GEMs. Pre-amplifiers are plugged 
      directly on its outer side and transmit the signal to the data 
      acquisition electronics.
\end{itemize}

@@

The GEM technology has been chosen for the flexibility of the GEM foils,
which can be easily used to produce a curved amplification surface. Also, GEMs 
are known to have relatively low spark rate~\cite{Bachmann:2000kj}, which is 
important when trying to detect highly ionizing slow nuclei that deposit large 
amount of energy. The GEMs for this RTPC are made from a Kapton insulator 
layer, 50~$\mu$m thick, sandwiched between two 5~$\mu$m copper 
layers\footnote{The GEM foils were produced by Tech-Etch, Inc.}. The mesh of 
each GEM layer is chemically etched with 50~$\mu$m diameter holes with 
double-conical shapes as illustrated in Figure~\ref{fig:GEMs}. The potential 
difference applied between the two copper layers of the GEM creates a very 
strong electric field in each hole leading to high ionization and 
amplification.

@@

The main structural improvement compared to the 2005 RTPC~\cite{Fenker:2008zz} 
was to obtain a better support structure for the GEM foils. In the previous version,
the RTPC was essencially composed of two independent half-cylinder separated
by their own structures. In these, the installation of the GEM was not very practical
and wrinkles were visibles on the GEM surfaces. To allow for a better installation 
and, at the same time, keep the mechanical structure out of the drift volume, 
the present design is based on GEM cylinder. This improves the acceptance of the
overall detector and reduces the material budget in the 
forward region, where we detect other particles in CLAS. We 
realized these cylinders by using fiber glass rings glued to 
each ends of the GEM foils to form self supporting pieces that could be 
installed independently in the RTPC after gluing and soldering operations. The 
rigidity of the GEM foils was enough for the structure to be self-supporting 
and only the upstream end of the cylinder was fixed to the main mechanical 
support structure. This design only left a light fiberglass ring in the 
downstream end, reducing to a minimum secondary interactions.

@@

%\begin{figure*}[tb]
%   \centering
%   \includegraphics[width=.75\textwidth]{ReadoutScheme.png}
%   \caption[]{Schematic representation of the RTPC readout system, showing
%   the different elements of the front end electronics.} 
%   \label{fig:ReadoutScheme}
%\end{figure*}

The readout system was an upgraded version of the original BoNuS RTPC system~\cite{Fenker:2008zz}, 
based ALICE-TPC front end electronic boards (FECs)~\cite{ALICE-FEE}. 
Each FEC hosted 128 channels, providing amplification and digitization of the input signals.  For each event, 100 samples/channel were digitized by the ALTRO ASIC~\cite{EsteveBosch:2003bj}, operated at 10 MHz sampling frequency. In order to reduce the data size, the system was operated in zero-suppression mode, keeping data from $N_{PRE}=3$ samples before threshold crossing on the rising edge to $N_{POST}=3$ samples after threshold crossing on the falling edge. The theshold level was set on each channel just above the noise level.

A new custom backplane was developed to connect FECs to the Readout Control Unit (RCU)~\cite{RCU}, allowing fast (200 MB/s) communication between the boards. The RCU board, used to distribute the trigger signal to different FECs and read data, was equipped with a fast optical link for data transferring to the main event builder. These features, together with the ``block-transfer'' readout mode making use of the FECs multi-event buffer, allowed to reach a signifantly higher readout-rate compared to the original BoNuS system.  During the 2009 run, the system was successfully 
operated with a DAQ rate of 3.1 kHz and a live time of $70 \%$, for a 
luminosity of about $10^{34}$~cm$^{-2}\cdot$s$^{-1}$ and a beam energy of 
6.064~GeV, to be compared with the 500 Hz obtained with the first BoNuS detector at similar run conditions.

@@

In this section we will detail the methods used to calibrate the drift speed,
drift paths and gains of the detector. Drift speed and paths were initially
calculated using the MAGBOLTZ~\cite{Biagi:1999nwa} program, then refined using
data to account for variations of the run conditions. We always assume 
cylindrical symmetry in the chamber for the calibration, such that none of
the parameters depend on the azimuthal angle $\phi$. The initial MAGBOLTZ
calibration was improved through several iterations of the
process described below, with each time an increasing number of tracks 
reconstructed in the RTPC. The figures presented in this section
are the ones obtained while performing the last iteration of this
calibration process.

\subsection{Drift Speed Parametrization}

After the ionization, the released electrons drift to the cylindrical detection 
plane under the effect of the electric field. The electrons released close to 
the cathode take the most time to reach the readout pads, and the cylindrical 
symmetry ensures that this maximum drift distance is the same for all tracks.  
We illustrate this in Figure~\ref{fig:RTPC_signals}, where we depict a typical 
$^{4}$He track. Due to the non-perfect experimental conditions, in particular 
possible contamination of our gas mixture, the drift speed changed during the 
three months long experimental run\footnote{We had to increase the gas flow 
during the experiment due to a small leak in the RTPC, which concurred to 
the large shift of drift time observed around run number 61600. While our 
gas system was kept slightly over atmospheric pressure to limit 
contamination from air or other external gases, it is likely that this leak was 
the source of modification of the drift speed.}. This can be calibrated by 
finding drift speed as function of experimental run-number. By measuring a 
reference time, we can therefore infer the drift speed of the electrons in the 
RTPC, by: $\textnormal{drift time} = \frac{\textnormal{drift path 
length}}{\textnormal{drift speed}}$.

@@

The drift path is the trajectory followed by the electrons released through 
ionization in the gas. We calculated them with
MAGBOLTZ~\cite{Biagi:1999nwa}, but it requires knowledge of the detector's 
geometry, gas mixture composition, and of course the electric and magnetic 
fields over the whole volume of the detector. We used this calculation as a 
first calibration, but, as can be seen with the drift speed, conditions in the 
chamber were changing over time. Moreover, the 4 $\mu$m foil used as a cathode 
is easily deformed, such that we expect the geometrical accuracy to be of few 
millimeters, directly impacting our knowledge of the electric field. These 
problems, already encountered for the BoNuS RTPC 
calibration~\cite{Fenker:2008zz}, motivated the acquisition of specific 
calibration runs. These were taken with a lower energy electron beam (1.204 and 
1.269 GeV) to enhance the cross section of the elastic scattering 
($e^{4}$He$\rightarrow e^{4}$He). In this process, the measurement of
the electron kinematics allows to calculate directly the kinematics of the Helium nucleus. 
By comparing the calculated momentum and angle of the recoil alpha particle to the 
measurement in the RTPC, we tuned the drift paths independently of our 
knowledge of the chamber's conditions.

@@

\subsection{Energy Loss Corrections}\label{sec_eloss}
Energy loss between the target and drift region was significant in the RTPC and 
necessitated a correction for optimal momentum reconstruction at the primary 
interaction vertex.  The dominant loss was in the 27~$\mu$m thick Kapton target 
wall, with significant contributions also from the pressurized target gas and 
the foils before reaching the drift region.  Corrections were developed based 
on GEANT4 simulations with the full RTPC geometry and parameterized in terms of 
recoil curvature in the drift region and polar angle, separetely for all recoil 
hypotheses ($p$, $d$, $^3$H, $^3$He, $^4$He).  At our average coherent $^4$He 
DVCS kinematics, energy losses were about 5 MeV, while for $e-^4$He elastic 
scattering losses were about 3 MeV, which corresponds to momentum corrections 
of 25\% and 15\%, respectively.

>>>>

\documentclass[preprint,5p]{elsarticle}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}  
\usepackage{textcomp}  
\usepackage{amssymb}  
\usepackage{hyperref}
\providecommand{\keywords}[1]{\textbf{\textit{Keywords:}} #1}
\hypersetup{colorlinks=true, urlcolor=blue, citecolor=blue}
\usepackage[displaymath]{lineno}
\linenumbers
\usepackage{dblfloatfix}    % To enable figures at the bottom of page
\begin{document}

@@

The original design of the RTPC was developed for the BoNuS 
experiment at Jefferson Lab which took data with CLAS in 
2005~\cite{Fenker:2008zz}. Significant improvements were made to the RTPC mechanical 
structure and fabrication technique that both increased the acceptance and 
reduced the amount of material in the path of the outgoing particles. 
Moreover, the data aquisition electronic was improved to increase the event 
readout rate. The enhanced design, 
used in the 2009 DVCS experiment, is described in section \ref{sec_design} of 
this paper. The data acquisition system is described in section \ref{sec_readout}, 
the calibration methods in section \ref{sec_calib} and the tracking algorithm in section
\ref{sec_tracking}. Finally, the overall performances
of the RTPC are described in section \ref{sec_perfor}.

@@

We detail here the different regions shown in Figure~\ref{fig:RTPC_1_4} 
starting from the beam line towards larger radius:\\
\begin{itemize}
  \item The 6~atm $^4$He target extends along the beamline forming the detector 
     central axis. It is a 6~mm diameter Kapton straw with a 27~\textmu{}m wall 
     of 292~mm length such that its entrance and exit 15~\textmu{}m aluminum windows 
     are placed outside of the detector volume. The detector and the target are 
     placed in the center of the solenoid, 64~cm upstream of the CLAS center.
   \item The first gas gap covers the radial range from 3~mm to 20~mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 \textmu{}m thick window made of grounded aluminized Mylar.
   \item The second gas gap region extends between 20~mm and 30~mm and is 
      filled with the gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether 
      (DME).  This region is surrounded by a 4~\textmu{}m thick window made of 
      aluminized Mylar set at $-4260$~V to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and 
      extends from the cathode to the first GEM, 60~mm away from the beam axis.  
      The average electric field in this region is perpendicular to the beam 
      and about 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66~mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600~V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charges 
      after they have been multiplied by the GEMs. Pre-amplifiers are plugged 
      directly on its outer side and transmit the signal to the data 
      acquisition electronics.
\end{itemize}

@@

The GEM technology has been chosen for the flexibility of the GEM foils,
which can be easily used to produce a curved amplification surface. Also, GEMs 
are known to have relatively low spark rate~\cite{Bachmann:2000kj}, which is 
important when trying to detect highly ionizing slow nuclei that deposit large 
amount of energy. The GEMs for this RTPC are made from a Kapton insulator 
layer, 50~\textmu{}m thick, sandwiched between two 5~\textmu{}m copper 
layers\footnote{The GEM foils were produced by Tech-Etch, Inc.}. The mesh of 
each GEM layer is chemically etched with 50~\textmu{}m diameter holes with 
double-conical shapes as illustrated in Figure~\ref{fig:GEMs}. The potential 
difference applied between the two copper layers of the GEM creates a very 
strong electric field in each hole leading to high ionization and 
amplification.

@@

The main structural improvement compared to the 2005 RTPC~\cite{Fenker:2008zz} 
was to obtain a better support structure for the GEM foils. In the previous version,
the RTPC was essencially composed of two independent half-cylinder separated
by their own structures. In this design, the installation of the GEMs was not very practical
and wrinkles were visibles on the GEM surfaces. To allow for a better installation 
and, at the same time, keep the mechanical structure out of the drift volume, 
the present design is based on self supporting GEM cylinder. We 
realized these cylinders by using fiber glass rings glued to 
each ends of the GEM foils in order to 
instal them independently in the RTPC after gluing and soldering operations. The 
rigidity of the GEM foils was enough for the structure to be self-supporting 
and only the upstream end of the cylinder was fixed to the main mechanical 
support structure. This design only left a light fiberglass ring in the 
downstream end, reducing to a minimum secondary interactions.

@@

The readout system was an upgraded version of the original BoNuS RTPC system~\cite{Fenker:2008zz}, 
based ALICE-TPC front end electronic boards (FECs)~\cite{ALICE-FEE}. 
Each FEC hosted 128 channels, providing amplification and digitization of the 
input signals. For each event, 100 samples/channel were digitized by the ALTRO 
ASIC~\cite{EsteveBosch:2003bj}, operated at 10 MHz sampling frequency. In order 
to reduce the data size, the system was operated in zero-suppression mode, 
keeping data from $N_{PRE}=3$ samples before threshold crossing on the rising 
edge to $N_{POST}=3$ samples after threshold crossing on the falling edge. 
The threshold level was set on each channel just above the noise level.

A new custom backplane was developed to connect FECs to the Readout Control 
Unit (RCU)~\cite{RCU}, allowing fast (200 MB/s) communication between the 
boards. The RCU board, used to distribute the trigger signal to different FECs 
and read data, was equipped with a fast optical link for data transferring to 
the main event builder. These features, together with the ``block-transfer'' 
readout mode making use of the FECs multi-event buffer, allowed to reach a 
signifantly higher readout-rate compared to the original BoNuS system. During 
the 2009 run, the system was successfully 
operated with a DAQ rate of 3.1 kHz and a live time of 70~\%, for a 
luminosity of about $10^{34}$~cm$^{-2}\cdot$s$^{-1}$ and a beam energy of 
6.064~GeV, to be compared with the 500 Hz obtained with the first BoNuS detector at similar run conditions.

@@

In this section we will detail the methods used to calibrate the drift time,
drift paths and gains of the detector. Drift time and paths were initially
calculated using the MAGBOLTZ~\cite{Biagi:1999nwa} program, then refined using
data to account for variations of the run conditions. We assume 
cylindrical symmetry in the chamber for the calibration of the drift parameters, 
such that do not depend on the azimuthal angle $\phi$. The initial MAGBOLTZ
calibration was improved through several iterations of the
data driven process described below, with each time an increasing number of tracks 
reconstructed in the RTPC. The figures presented in this section
are the ones obtained while performing the last iteration of this
calibration process.

\subsection{Maximum Drift Time Parametrization}

%To be rewritten!!!
After the ionization, the released electrons drift to the cylindrical detection 
plane under the effect of the electric field. The electrons released close to 
the cathode take the most time to reach the readout pads, the cylindrical 
symmetry indicates that this maximum drift distance is the same for all tracks.  
We illustrate this in Figure~\ref{fig:RTPC_signals}, where we depict a typical 
$^{4}$He track. Due to the non-perfect experimental conditions, in particular 
possible contamination of our gas mixture, the drift speed changed during the 
three months long experimental run\footnote{We had to increase the gas flow 
during the experiment due to a small leak in the RTPC, which concurred to 
the large shift of drift time observed around run number 61600. While our 
gas system was kept slightly over atmospheric pressure to limit 
contamination from air or other external gases, it is likely that this leak was 
the source of modification of the drift speed.}. This can be calibrated by 
finding drift speed as function of experimental run-number. By measuring a 
reference time, we can therefore infer the drift speed of the electrons in the 
RTPC, by: $\textnormal{drift time} = \frac{\textnormal{drift path 
length}}{\textnormal{drift speed}}$.

@@

The drift path is the trajectory followed by the electrons released through 
ionization in the gas. We calculated them with
MAGBOLTZ~\cite{Biagi:1999nwa}, but it requires knowledge of the detector's 
geometry, gas mixture composition, and of course the electric and magnetic 
fields over the whole volume of the detector. We used this calculation as a 
first calibration, but, as can be seen with the drift speed, conditions in the 
chamber were changing over time. Moreover, the 4 \textmu{}m foil used as a cathode 
is easily deformed, such that we expect the geometrical accuracy to be of few 
millimeters, directly impacting our knowledge of the electric field. These 
problems, already encountered for the BoNuS RTPC 
calibration~\cite{Fenker:2008zz}, motivated the acquisition of specific 
calibration runs. These were taken with a lower energy electron beam (1.204 and 
1.269 GeV) to enhance the cross section of the elastic scattering 
($e^{4}$He$\rightarrow e^{4}$He). In this process, the measurement of
the electron kinematics allows to calculate directly the kinematics of the Helium nucleus. 
By comparing the calculated momentum and angle of the recoil alpha particle to the 
measurement in the RTPC, we tuned the drift paths independently of our 
knowledge of the chamber's conditions.

@@

\subsection{Energy Loss Corrections}\label{sec_eloss}
Energy loss between the target and drift region was significant in the RTPC and 
necessitated a correction for optimal momentum reconstruction at the primary 
interaction vertex.  The dominant loss was in the 27~\textmu{}m thick Kapton target 
wall, with significant contributions also from the pressurized target gas and 
the foils before reaching the drift region.  Corrections were developed based 
on GEANT4 simulations with the full RTPC geometry and parameterized in terms of 
recoil curvature in the drift region and polar angle, separetely for all recoil 
hypotheses ($p$, $d$, $^3$H, $^3$He, $^4$He).  At our average coherent $^4$He 
DVCS kinematics, energy losses were about 5 MeV, while for $e-^4$He elastic 
scattering losses were about 3 MeV, which corresponds to momentum corrections 
of 25\% and 15\%, respectively.
